\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}

### 2013-06-26 paper.tex
Major changes in line with Andy and Saras requests.

The computational prediction of mutational effects on protein--protein interactions remains a challenging problem. Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure. For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}. However, such methods calculate the relative contribution to overall energy for each amino acid in a protein. They do not probe the interface directly, but rather calculate the total energy with each substitution \citep{Grant2011,Kortemme2004}. Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}. Unfortunately, such machine-learning approaches often suffer in novel applications, for which available training sets are small or non-existent. As such, these methods are poorly suited for most host-virus protein--protein systems. By contrast, first principles methods offer the chance to forgo training. However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model that phases directly from one three-dimensional protein model to another with no intermediate transition states \citep{Gilson1997,Lu2004}. Thus, they can quickly lose physiological relevance as the two proteins become more dissimilar and as a result are largely limited to testing only sequential point mutants.

Here we propose a new method to probe mutational effects in protein--protein interactions directly, by pulling two proteins apart that are initially bound in a complex. We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place. Then, we measure the force required for dissociation. As a proxy for relative binding affinity, we use the maximum applied force (MAF) required for separation and the area under the force-versus-distance curve (AUC).

We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}. Machupo is a negative-sense RNA virus of the arenavirus family \citep{Charrel2003}. Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}. Members of the arenavirus family include the lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the guanarito, junin, and machupo viruses endemic to rodents in South America \citep{Charrel2003}. The South American arenaviruses typically infect humans after rodent contamination and cause a devastating hemorrhagic fever with very high mortality \citep{Charrel2003}.

The hTfR1 is the primary receptor used by MACV for binding its host cell prior to infection. The hTfR1 protein contains three extracellular domains with the two more basilar domains serving most of the endogenous, transferrin-binding function \citep{Abraham2010,Rad20112}. Cellular entry is initiated by GP1 binding to the apical domain of hTfR1. Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of viral infectivity \citep{Rad20111,Rad20112}. The co-crystal structure shows that a 15 amino acid flexible loop on GP1 is responsible for viral binding \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}. Furthermore, the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain. Experimental alanine-scanning and whole cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.

We applied our computational method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes. At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity is often not as simple as the static structure suggests. For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1. According to our simulations, this network likely does not affect binding affinity. Our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in host-virus protein--protein interfaces.

@@

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. To reduce the noise, we averaged each point with a 75 point interpolation window. The window ensured that the averaged MAF was never cut off near the beginning of the simulation due to lost time steps (caused by averaging). The two primary descriptive statistics we used were maximum interpolated applied force (MAF) and total area under the interpolated curve (AUC). We used the pairwise.t.test function with the FDR p-value correction to test for significance. The ggplot \citep{ggplot} package was used to generate most of the figures.

@@

\subsection*{The GP1/hTfR1 system}

The GP1/hTfR1 interface marks a particularly important and useful test system. There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV. Virtually all of the important sites have been mapped by \textit{in vitro} flow-cytometry based infectivity assays. The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals). In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \citep{Grant2011,Kortemme2004} only marginally useful. The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.

@@

\subsection*{Steered molecular dynamics simulations}
To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions. First, we generated a physically realistic simulation world around the protein complex. The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions. Next, we carried out successive rounds of energy minimization with a fixed $\alpha$-carbon backbone, followed by heating, and backbone release. Last, we performed a short round of all-atom molecular dynamics with minimal restraints \citep{Cuendet2008}. We used the final state from each equilibrated system to restart another MD simulation. For SMD, at the centroid of several atoms a force was applied to a single point in GP1. By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging. The latter approach was used for statistical analysis. In addition, to minimize the computational time in unproductive molecular dynamics, we shortened the simulation time such that the interaction just barely dropped to zero near the end of each run. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. The supplementary movie visually illustrates the separation distance between peptide domains. The quantities MAF and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

Before systematically applying SMD to the GP1/hTfR1 interaction, we needed to ensure the method was sufficiently sensitive to distinguish between relatively minor point mutations. While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape. For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach. Several residues on both proteins appear to mediate putative hydrogen-bonding interactions. We selected three sites for alanine substitutions: Tyr211 and Asn348 in the hTfR1 and vArg111 in MACV. For notation purposes, the viral site is always referred to with a preceding `v'. According to the static complex \citep{Abraham2010}, Tyr211 and vArg111 are involved in one hydrogen-bonding network and Asn348 is involved in a second, distinct hydrogen-bonding network.

We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the MAF and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:pairwise_differences}), while the other two did not (Table~\ref{tab:pairwise_differences}). From these results, we concluded that our method can indeed successfully distinguish individual mutants from WT, and that alanine substitutions at Asn348 and vArg111 may have little effect on binding affinity (see below for further discussion). When considering additional mutants (also discussed below), we found that MAF was generally sufficient to distinguish mutants (Tables~\ref{tab:pairwise_differences} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}). In general, however, MAF seemed to be the more sensitive statistic than AUC.

\subsection*{Comparative analysis of the GP1/hTfR1 interface}

Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. We made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause reduced binding. Then, we compared the size of those differences to reduced binding single mutants (Figure~\ref{fig:sensitivity}).

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_statistics}). The largest single-mutant impact was seen for a mutation to alanine at site Tyr211 in the hTfR1 ($ p < 0.001 $).  Several mutations at this site have proven capable of causing loss-of-infection, according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 in GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure \citep{Abraham2010}. In agreement with the static structure, our molecular dynamics simulations show Tyr211 to be involved in a large hydrogen-bonding network (results not shown).

Although Tyr211Ala appears to have a large impact on binding affinity, the single mutant provided little underlying biochemical rationale. Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for loss of binding affinity. In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding. We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}. Both mutations proved incapable of causing reduced binding affinity in our simulations. In fact, they caused a significant increase in relative binding affinity (Figure~\ref{fig:sensitivity} and Table~\ref{tab:MAF_p}). This outcome is not surprising, if we consider that Tyr211 participates in a bonded network rather than independently. Combined with some biochemical intuition, these results suggest that mutations to other polar residues would likely not affect the dynamics of binding directly.

We also simulated several point mutations at Asn348 in the hTfR1. Asn348Lys is reported in the literature to cause reduced infectivity \textit{in vivo} \citep{Rad2008,Abraham2010}. Further, in the crystal structure two residues on GP1 and at least three on hTfR1, including Asn348, are involved in a fairly extensive hydrogen-bonding network. As discussed above, the alanine mutation at this site showed no significant difference in MAF or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp. (For both of these mutations, however, mean MAF and mean AUC was lower than for WT. See Table~\ref{tab:summary_statistics}.)  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. These results suggest that, in contrast to Tyr211, the critical difference at position 348 in hTfR1 is size exclusion rather than the destruction of a hydrogen-bonding network. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:pairwise_differences}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to MAF (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

Last, we further analyzed a single mutation in GP1, vArg111Ala. It has been reported that vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}. As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either MAF or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}). This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity. To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala. This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:pairwise_differences}). Thus, vArg111Ala in the GP1 appeared capable of suppressing the loss-of-binding caused by Tyr211Ala. This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction leads to comparable binding affinity.

\section*{Discussion}
We have applied a method utilizing steering forces in all-atom molecular dynamics simulations to evaluate the effects of mutations at the GP1/hTfR1 interface. We modeled mutations at several sites in the GP1/hTfR1 interface, and verified that our computational protocol was sensitive enough to distinguish point mutants in hTfR1. Further, we identified two test statistics, MAF and AUC, that can be used as proxies for binding affinity. We systematically tested several point mutations to understand their contribution to the binding interaction. In the case of Asn348Lys, we have shown that the static structure provides little insight into why this mutation causes loss-of-infectivity \textit{in vivo}. While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, loss of binding affinity at that site is actually caused by size and charge restriction. Further, we have shown that the cause of diminished infectivity in the viral single-mutant vArg111Ala may not be due to binding-energy differences. We also found that a negatively polar residue at site 211 in hTfR1 is absolutely critical for a tight binding interaction. Any non-polar mutation at Tyr211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection. However, our data suggest that the viral mutation vArg111Ala may restore the infective potential of MACV that was previously destroyed by a Tyr211Ala hTfR1 mutation.

@@

Our findings rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}. We found that some substitutions at positions 211 and 348 did negatively affect the strength of viral binding. However, the computational data suggest that the reason and nature of the effects at these two sites are very different. Congruent with infectivity data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding. Tyr211Ala alone produces such a large decrease in affinity that it can be easily distinguished from the WT complex. Structural analysis reveals that the biochemical cause involves the destruction of a critical hydrogen-bonding network with Ser113 and vArg111 in GP1. Further, Tyr211Asp and Tyr211Thr were easily resolved from the WT complex, but they appear to cause an increase in affinity. As a result, we predict experimental structural analysis will show that Tyr211Asp and Tyr211Thr mutations cause a relatively dramatic tertiary structure deviation. Given that our model starts from an associated complex, it is not surprising that Tyr211Asp would cause increased rather than decreased affinity. Aspartate is charged rather than polar at physiological pH and its shorter side chain means the proteins would be pulled in closer apposition. Nevertheless, our results show Tyr211, or some other amino acid that is capable of hydrogen bonding, is necessary for a stable hydrogen-bonding network leading to efficient binding. Moreover, the cause of decreased infectivity seen in rodent populations with the Tyr211Asp mutation is not a simple change in binding affinity.

Previous work has shown that the hTfR1 mutation Asn348Lys is capable of producing decreased infectivity \emph{in vitro} \citep{Rad2008,Rad20111}. Moreover, the molecular model based on x-ray diffraction data suggests destruction of a second hydrogen-bonding network as the cause \citep{Abraham2010}. In our simulations no individual point mutation at Asn348 showed a significantly different affinity from the WT complex. Moreover, the mutation Asn348Ala resulted in an average MAF that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}). Thus, the putative hydrogen-bonding network involving Asn348 is probably a crystallization artifact. Even when paired with the highly destructive Tyr211Ala mutation, Asn348Ala was refractory to changes in binding affinity. By contrast, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Moreover, although Asn348Trp and Asn348Lys individually failed to be distinguishable from WT hTfR1, they both caused a mean reduction in MAF and produced larger than expected synergistic effects. This data suggests that site 348 is size and charge restricted. Any mutation to a very large or positively charged side chain will likely produce decreased infectivity.

>>>>

The computational prediction of mutational effects on protein--protein interactions remains a challenging problem. Several methods are available to perform an energy difference calculation from an experimentally determined co-crystal structure. For example, computational alanine-scanning can be performed rapidly, with relatively low computational cost \citep{Grant2011,Kortemme2004}. However, such methods calculate the relative contribution to overall energy for each amino acid in a protein. They do not probe the interface directly, but rather calculate the total energy with each substitution \citep{Grant2011,Kortemme2004}. Alternative approaches have been developed using machine learning, training coefficients in a weighted equation containing geometric and energetic parameters \citep{Vreven2011,Vreven2012,Bajaj2011,Hwang2010}. Unfortunately, such machine-learning approaches often suffer in novel applications, for which available training sets are small or non-existent. As such, these methods are poorly suited for most host-virus protein--protein systems. By contrast, first principles methods offer the chance to forgo training. However, the currently available methods such as free energy perturbation (FEP) and thermodynamic integration (TI) rely on a two state model (where one state may be wild-type and the other may be a mutant) that transitions directly from one three-dimensional protein model to another with no intermediate steps \citep{Gilson1997,Lu2004}. As a result, the accuracy of the measurement suffers as the two proteins become increasingly dissimilar; therefore, these techniques are largely limited to point mutant comparisons.

Here we propose a new method to probe mutational effects in protein--protein interactions directly, by pulling two proteins apart that are initially bound in a complex. We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place. Then, we measure the force required for dissociation. As a proxy for relative binding affinity, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC).

We applied our method to interrogate the interaction between machupo virus (MACV) spike glycoprotein (GP1) and the human transferrin receptor (hTfR1) \citep{Abraham2010,Charrel2003}. Machupo is an ambisense RNA virus of the arenavirus family \citep{Charrel2003}. Worldwide, arenaviruses represent a significant source of emerging zoonotic diseases for the human population \citep{Charrel2003}. Members of the arenavirus family include the Lassa fever virus endemic to West Africa, the lymphochoriomeningitis virus (LCMV) endemic to rodents in several areas of the United States, and the Guanarito, Junin, and Machupo viruses endemic to rodents in South America \citep{Charrel2003}. The South American arenaviruses typically infect humans after rodent contamination and can cause a devastating hemorrhagic fever with high mortality \citep{Charrel2003}.

The hTfR1 is the primary receptor used by MACV for binding its host cell prior to infection. The primary role of hTfR1 $in~vivo$ is to bind transferrin for cellular iron uptake. The hTfR1 protein contains three extracellular domains: two basilar domains and an apical domain. The two basilar domains serve most of the transferrin-binding function \citep{Abraham2010,Rad20112}. Viral entry is initiated by GP1 binding to the apical domain of hTfR1. Previous work has indicated that the GP1/hTfR1 binding interaction is the primary determinant of MACV host range variation \citep{Rad20111,Rad20112}. The co-crystal structure shows the high affinity interaction between GP1 and hTfR1 forces the normally flexible loop in the apical domain of hTfR1 into a rigid beta-pleated sheet domain. For GP1, several extended loops in mediate binding to hTfR1 \citep{Abraham2010,Rad20112}, and many of the interface interactions are mediated by extensive hydrogen-bonding networks \citep{Abraham2010}. Experimental alanine-scanning and whole cell infectivity assays have identified several sites in both GP1 and hTfR1 that are probably critical for establishing infection \citep{Rad20111,Rad20112}.

We applied our computational method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes. At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity is often not as simple as the static structure suggests. For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1. According to our simulations, this network likely does not affect binding affinity directly. Our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in protein--protein interfaces.

@@

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. To reduce the noise, we averaged each point with a 75 point interpolation window. The window ensured that the averaged maximum applied force was never cut off near the beginning of the simulation due to lost time steps (caused by averaging). The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC). We used the pairwise.t.test function with the FDR p-value correction to test for significance. The ggplot \citep{ggplot} package was used to generate most of the figures.

@@

\subsection*{The GP1/hTfR1 system}
The GP1/hTfR1 interface marks a particularly important and useful test system. There are several sites on both the human and viral protein known to affect the infectivity phenotype of MACV. Virtually all of the important sites have been mapped by \textit{in vitro} flow-cytometry based entry assays. The GP1/hTfR1 interface appears not to be dominated by one particular type of interaction (electrostatics, hydrogen-bonding, or van der Waals). In addition, much of the binding domain on hTfR1 is on a flexible loop and undergoes a large structural rearrangement upon binding, making many other computational techniques \citep{Grant2011,Kortemme2004} only marginally useful. The complex nature of this interface represents a particularly difficult challenge for traditional computational analysis.

@@

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants}). Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}. In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1. In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}. Finally, an alanine mutation at site 111 in GP1 has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}. For notation purposes, the viral site is always referred to with a preceding `v'.

\subsection*{Steered molecular dynamics simulations}
To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions. First, we generated a physically realistic simulation world around the protein complex. The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions. Next, we carried out successive rounds of energy minimization with a fixed $\alpha$-carbon backbone, followed by heating, and backbone release. Last, we performed a short round of all-atom molecular dynamics with minimal restraints \citep{Cuendet2008}. We used the final state from each equilibrated system to restart another MD simulation. For SMD, at the centroid of several atoms a force was applied to a single point in GP1. By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging. The latter approach was used for statistical analysis. In addition, to minimize the computational time in unproductive molecular dynamics, we shortened the simulation time such that the interaction just barely dropped to zero near the end of each run. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. The supplementary movie visually illustrates the separation distance between peptide domains. The quantities maximum applied force and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

Before systematically applying SMD to the GP1/hTfR1 interaction, we needed to ensure the method was sufficiently sensitive to distinguish between relatively minor point mutations. While SMD has been applied previously to measure the binding energy of high-affinity T-cell receptor interactions \citep{Cuendet2008,Cuendet2011}, to our knowledge the method has never been used previously to parse small differences in a complex energy landscape. For this initial sensitivity analysis, we tested alanine substitutions congruent with the traditional experimental and computational approach.

We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:pairwise_differences}), while the other two did not (Table~\ref{tab:pairwise_differences}). From these results, we concluded that our method can indeed successfully distinguish individual mutants from WT, and that alanine substitutions at Asn348 and vArg111 may have little effect on binding affinity (see below for further discussion). When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:pairwise_differences} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}). In general, however, maximum applied force seemed to be the more sensitive statistic than AUC.

\subsection*{Comparative analysis of the GP1/hTfR1 interface}
Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. Seen in Table~\ref{tab:summary_mutant}), we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause reduced binding. Then, we compared the size of those differences to reduced binding single mutants (Figure~\ref{fig:sensitivity}).

Although Tyr211Ala appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism. Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for loss of binding affinity. In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding. We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}. Both mutations proved incapable of causing reduced binding affinity in our simulations. In fact, they caused a significant increase in relative binding affinity (Figure~\ref{fig:sensitivity} and Table~\ref{tab:MAF_p}).

We also simulated several point mutations at Asn348 in the hTfR1. As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp. (For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT. See Table~\ref{tab:summary_statistics}.)  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. These results suggest that, in contrast to Tyr211, the critical difference at position 348 in hTfR1 is size exclusion rather than the destruction of a hydrogen-bonding network. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:pairwise_differences}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to maximum applied force (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

Last, we further analyzed a single mutation in GP1, vArg111Ala. As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either maximum applied force or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}). This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity. To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala. This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:pairwise_differences}). Thus, vArg111Ala in the GP1 appeared capable of suppressing the loss-of-binding caused by Tyr211Ala. This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction leads to comparable binding affinity.

\section*{Discussion}

We have applied a method utilizing steering forces in all-atom molecular dynamics simulations to evaluate the effects of mutations at the GP1/hTfR1 interface. We modeled mutations at several sites in the GP1/hTfR1 interface, and verified that our computational protocol was sensitive enough to distinguish point mutants in hTfR1. Further, we identified two test statistics, maximum applied force and AUC, that can be used as proxies for binding affinity. We systematically tested several point mutations to understand their contribution to the binding interaction. In the case of Asn348Lys, we have shown that the static structure provides little insight into why this mutation causes loss-of-infectivity \textit{in vivo}. While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, loss of binding affinity at that site is actually caused by size and charge restriction. Further, we have shown that the cause of diminished infectivity in the viral single-mutant vArg111Ala may not be due to binding-energy differences. We also found that a negatively polar residue at site 211 in hTfR1 is absolutely critical for a tight binding interaction. Any non-polar mutation at Tyr211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection. However, our data suggest that the viral mutation vArg111Ala may restore the infective potential of MACV that was previously destroyed by a Tyr211Ala hTfR1 mutation.

@@

Our findings rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}. First, we found that some substitutions at positions 211 and 348 did negatively affect the strength of receptor binding. However, the computational data suggest that the reason and nature of the effects at these two sites are very different. Congruent with entry data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding. Second, no individual point mutation at Asn348 showed a significantly different affinity from the WT complex. However, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations. Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone. As mutations at hTfR1 site 211 are known to reduce infectivity \emph{in vitro} and \emph{in vivo}, it is possible that vArg111 in GP1 can serve as a site for compensatory MACV mutations.

%Structural analysis reveals that the biochemical cause involves the destruction of a critical hydrogen-bonding network with Ser113 and vArg111 in GP1. Further, Tyr211Asp and Tyr211Thr were easily resolved from the WT complex, but they appear to cause an increase in affinity. As a result, we predict experimental structural analysis will show that Tyr211Asp and Tyr211Thr mutations cause a relatively dramatic tertiary structure deviation. Given that our model starts from an associated complex, it is not surprising that Tyr211Asp would cause increased rather than decreased affinity. Aspartate is charged rather than polar at physiological pH and its shorter side chain means the proteins would be pulled in closer apposition. Nevertheless, our results show Tyr211, or some other amino acid that is capable of hydrogen bonding, is necessary for a stable hydrogen-bonding network leading to efficient binding. Moreover, the cause of decreased infectivity seen in rodent populations with the Tyr211Asp mutation is not a simple change in binding affinity.

%Previous work has shown that the hTfR1 mutation Asn348Lys is capable of producing decreased infectivity \emph{in vitro} \citep{Rad2008,Rad20111}. Moreover, the molecular model based on x-ray diffraction data suggests destruction of a second hydrogen-bonding network as the cause \citep{Abraham2010}. In our simulations no individual point mutation at Asn348 showed a significantly different affinity from the WT complex. Moreover, the mutation Asn348Ala resulted in an average maximum applied force that was virtually identical to WT hTfR1 (Table~\ref{tab:pairwise_differences}). Even when paired with the highly destructive Tyr211Ala mutation, Asn348Ala was refractory to changes in binding affinity. By contrast, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Moreover, although Asn348Trp and Asn348Lys individually failed to be distinguishable from WT hTfR1, they both caused a mean reduction in maximum applied force and produced larger than expected synergistic effects. This data suggests that site 348 is size and charge restricted. Any mutation to a very large or positively charged side chain will likely produce decreased infectivity.
